---
title: "Semester_Project"
format: html
---

# Introduction 
....

# Pre-Processing 
•	Daten anreichern: 
    	Oder einfach mit Segmenten arbeiten? Sessile and moving unterscheiden?
  o	Unterscheidung Regen/kein Regen (Wetterstationen Voronoi)
  o	Uni Zugehörigkeit anreichern? 


## Packages 
```{r}
# clear space 
rm(list=ls())

# load packages 
library("readr")
library("dplyr")
library("ggplot2")
library("sf")
library("terra")
library("tmap")
library("gitcreds")
library("dplyr")
library("SimilarityMeasures")
library("lubridate")
library("plotly")
```

## Load Data 
```{r}
# load all posmo data --------------------------------------- 
posmo1 <- read_delim("posmo_data/steinm05_posmo_2023_04_10_2023_05_15.csv")
posmo2 <- read_delim("posmo_data/scheimir_posmo_2023-01-01T00_00_00+01_00-2023-05-15T23_59_59+02_00.csv")
posmo3 <- read_delim("posmo_data/a77bd18b-2615-4b03-802b-879bf3c93798_2023-01-01T00-00-00_2023-05-13T23-59-59.csv")
posmo4 <- read_delim("posmo_data/posmo - BuJa.csv")
posmo5 <- read_delim("posmo_data/posmo.csv")

# gps1 <- read_delim("posmo_data/gps_bike_walk_train.csv", ";")
```

## Filter data/ create workable dataframe
```{r}
# User anschreiben/umbenennen ---------------------------------------------
# change user_id to names, so that we can read better who is who 
posmo1 <- posmo1 |>  
  mutate(dataset = "posmo1")

posmo2 <- posmo2 |>  
  mutate(dataset = "posmo2")

posmo3 <- posmo3 |>  
  mutate(dataset = "posmo3")

posmo4 <- posmo4 |>  
  mutate(dataset = "posmo4")

posmo5 <- posmo5 |>  
  mutate(dataset = "posmo5")


# transform posmo data into an sf object ----------------------------------
posmo1 <- st_as_sf(posmo1, coords = c("lon_x","lat_y"), crs = 4326) |>
  st_transform(2056)

posmo2 <- st_as_sf(posmo2, coords = c("lon_x","lat_y"), crs = 4326) |>
  st_transform(2056)

posmo3 <- st_as_sf(posmo3, coords = c("lon_x","lat_y"), crs = 4326) |>
  st_transform(2056)

posmo4 <- st_as_sf(posmo4, coords = c("lon_x","lat_y"), crs = 4326) |>
  st_transform(2056)

posmo5 <- st_as_sf(posmo5, coords = c("lon_x","lat_y"), crs = 4326) |>
  st_transform(2056)

# bind all data together, create a single dataframe -------------------------
posmo <- rbind(posmo1, posmo2, posmo3, posmo4, posmo5)

# Datetime in POSIXct format umwandeln -------------------------------------
# Sonstige Datenformate anpassen bei Bedarf --------------------------------
posmo <- posmo |>
  mutate(datetime = as.POSIXct(datetime),
         weekday = as.factor(weekday),
         transport_mode = as.factor(transport_mode),
         dataset = as.factor(dataset))
```

### remove outliers 
```{r}
# plot data to look at areas needed to be covered for weather stations 
tmap_mode("view")

# tm_shape(posmo) +
 # tm_dots()

# look at single posmo data to determine outliers 
# outlier - norway
tm_shape(posmo1) +
  tm_dots()

# no outlier
tm_shape(posmo2) +
  tm_dots()

# no outlier
tm_shape(posmo3) +
  tm_dots()

# no outlier 
tm_shape(posmo4) +
  tm_dots()

# outlier - france 
tm_shape(posmo5) +
  tm_dots()

# delete trajectory data outside of switzerland (= delete outliers; norway & france) ------------------------------------------------------------------
# load swiss boundary 
swiss <- read_sf("data/schweiz.gpkg")

# cut layer with swiss boarder
result_layer <- st_intersection(posmo, swiss)

# now outliers are out 
tm_shape(result_layer) +
 tm_dots()
```

## Add convenience variables 
```{r}
# Unterscheidung zwischen Wochentag und Wochenende (Sa-So: Weekend, Mo-Fr: Workday) -----------------------------------------------------------------

# geht noch nicht 
posmo <- posmo |>
  mutate(day_week = ifelse(weekday == "Sat" & weekday == "Sun", "weekend", "weekday")) 

# Unterscheidung Tag, Nacht (Tag: 06:00 – 22:00, Nacht: 22:00 – 06:00?) --------------

# geht noch nicht
posmo <- posmo |>
  mutate(time = format(datetime, format = "%H:%M:%S"), # Add convenience variable with only time 
         day_night = ifelse(time >= "22:00:00" & time <= "06:00:00", "night", "day")) # divide based on time into night and day
```

# Research Questions 
Does the weather condition (e.g. rain, sunshine) impact spatio-temporal patterns of students by foot and/or bicycle?

o	Group by movement (nicht bewegt, bewegt), Tag/nacht, Datum, Wettercondition, User_id
o	Mittlere Distanz gesamt ausrechnen 
o	Mittlere Distanz pro Travelmode 
o	Mittlere Distanz Fussverkehr 
o	Mittlere Distanz Veloverkehr 

```{r}

```

Does the day of the week (weekend vs. workday) have an impact on spatio-temporal movement patterns? 

o	Group by movement (nicht bewegt, bewegt), Tag/nacht, Datum, Tag, User_id
o	Mittlere Distanz gesamt ausrechnen 
o	Mittlere Distanz pro Travelmode 

```{r}

```

Do students of different universities (ZHAW or UZH) show different spatio-temporal movement patterns?
```{r}

```

# Visualisations 
Wetterfrage: 
•	Boxplot 
o	pro User und Wettercondition 
o	pro Wettercondition 
o	Pro travel mode 
o	Pro Fussverkehr, Veloverkehr

