---
title: "Semester_Project"
format: html
---

# Pre-Processing 
•	Daten bereinigen 
  o	User anschreiben/umbenennen, 
  o	Outliers: Norwegen entfernen, sonstige Outliers entfernen 
  o	Datetime in POSIXct format umwandeln 
  o	Sonstige Datenformate anpassen bei Bedarf
•	Daten anreichern: 
  o	Unterscheidung Tag, Nacht (Tag: 06:00 – 22:00, Nacht: 22:00 – 06:00?)
    	Oder einfach mit Segmenten arbeiten? Sessile and moving unterscheiden?
  o	Unterscheidung Regen/kein Regen (Wetterstationen Voronoi)
  o	Uni Zugehörigkeit anreichern? 
  o	Unterscheidung zwischen Wochentag und Wochenende (Sa-So: Weekend, Mo-Fr: Workday)

## Packages 
```{r}

```

## Load Data 
```{r}

```

# Research Questions 
Does the weather condition (e.g. rain, sunshine) impact spatio-temporal patterns of students by foot and/or bicycle?
```{r}

```

Does the day of the week (weekend vs. workday) have an impact on spatio-temporal movement patterns? 
```{r}

```

Do students of different universities (ZHAW or UZH) show different spatio-temporal movement patterns?
```{r}

```


