---
title: "Effects of three environmental factors on spatio-temporal movement patterns of students"
subtitle: "Project for Patterns and Trends in Environmental Data - Computational Movement Analysis"
author: "Mirjam Scheib & Miriam Steinhauer"
date: "07/02/2023"
format:
  html:
    code-fold: true
    code-tools: true
    code-line-numbers: true
highlight-style: pygments
table-of-contents: true
number-sections: true
execute:
  echo: true
  message: false
  eval: true
  warning: false
editor: 
  markdown: 
    wrap: 72
---

# BACKGROUND AND RESEARCH QUESTIONS

It is known that several factors like the weather condition (Brum-Bastos
et al., 2018; Guo et al., 2022), the day of the week and time of the day
(Liu et al., 2020; Sathishkumar et al., 2020) influence spatio-temporal
movement patterns of people. There are several studies investigating
movement patterns of people in urban areas with the aim to improve the
distribution of facilities and the provisioning of transportation
services, as well as to manage traffic peaks (Kyaing et al., 2017; Liu
et al., 2020). It was shown that travel behaviour is significantly
impacted by age, income and life stage influencing travelling behaviour
in different ways (Kattiyapornpong & Miller, 2009). Because people with
different socio-demographic variables follow distinct spatio-temporal
movement patterns, we investigate the travelling behaviour of 11
students in this project work. The canton of Zurich accommodates the
most students in Switzerland, with numbers continuing to rise (Bundesamt
für Statistik, 2023; Medienmitteilung Kanton ZH, 2021). Additionally,
students and apprentices (from the age of 15 years) make up 39 % of the
public transport commuter mass, which is why detailed knowledge of
spatio-temporal patterns of students could help to manage public
transport infrastructure (Bundesamt für Statistik, 2021). Therefore, we
want to investigate the influence of three environmental factors on
spatio-temporal movement patterns of students, aiming to answer the
following research questions:

1.  Does the day of the week (weekend vs. weekday) have an influence on
    the spatio-temporal movement patterns of students?
2.  Does the time of the day compared between weekend and weekday have
    an influence on the spatio-temporal movement patterns of students?
3.  Does intense precipitation have an impact on spatio-temporal
    movement patterns of students?

# DATA & METHODS

## Data

Primary data used in this project work consisted of trajectory data from
11 students collected with the Posmo App containing the following
attributes:

-   **user_id:** entails the individual ID of the user (= student)
    (type: character)
-   **datetime:** date and time when a position of a user is tracked
    (type: datetime)
-   **weekday:** abbreviated name of the weekday (e.g. Mon = Monday)
    (type: character)
-   **place_name:** names of a place, in which a user is at a specific
    time (type: character)
-   **transport_mode:** the type of transport used by a user (e.g. car,
    bicycle, foot, train, other....) (type: character)
-   **lon_x/lat_y:** coordinates of the user at a specific time (type:
    numeric)

Additionally, precipitation data consisting of measurement every 10
minutes from 84 stations in the canton of Zurich and bordering cantons
(SZ, AG, ZG, TG, SG) was used to investigate the influence of
precipitation on spatio-temporal movement patterns of students.

## Pre-processing

First we reviewed data consistency by creating a point plot, where every
tracking point in time (x-axis) was displayed per user (y-axis). Five
students needed to be eliminated due to insufficient tracking point
coverage in time (Fig. 1). Additionally, the dataset was cut into a
specific timeframe (28.04.2023 - 09.05.2023) covering 11 days to ensure
sufficient overlap of tracking points between users (Fig. 2).

![Figure 1: Point plot of every tracking point in time (x-axis)
displayed per student (y-axis). Posmo4, posmo7, posmo8, posmo9 and posmo
10 needed to be eliminated due to insufficient tracking point coverage
in time.](Cut_Time_Before.png)

![Figure 2: Point plot of every tracking point in time (x-axis)
displayed per student (y-axis). The dataset consisting of 6 students was
cut into a timespan covering 11 days to ensure sufficient tracking point
coverage for every student.](Cut_Time_After.png)

Additionally, travel modes consisting of "Airplane", "Funicular" and
"Other1" were removed, as they presented either too specific or
unspecific travel modes to be of interest in the subsequent analysis.
Furthermore, the annotation of all convenience variables are considered
crucial pre-processing steps for the subsequent analysis. First, it was
discriminated between weekend (Sa -- So) and weekdays (Mo -- Fr).
Additionally, to discriminate between days with high and low/no
precipitation we calculated the nearest weather station for every
trajectory point and annotated the precipitation data at the matching
time to our dataset. For this we created a new column in the posmo
dataset, to create a time join key matching the 10 minutes interval of
the weather data. We considered a precipitation of \> 20 mm per 24 hours
as a day with intense precipitation, as MeteoSwiss (2023) defines a
precipitation from 10 -- 30 mm per 24 hours as high intensity
precipitation.

To investigate spatio-temporal movement patterns of students in concern
to environmental variables, we discriminated between static and movement
segments and assigned to every moving sequence a segment ID. From this
we calculated distance, speed and duration for every segment.

All pre-processing steps can be accessed in a separate quarto-file
(Pre_Processing.qmd) in our github respiratory under the following link:
https://github.com/mirjamscheib/Semester_Project.git

## Analysis

The analysis of the research questions was carried out by creating
summary tables and boxplots to investigate distance, speed and duration
of different travel modes comparing weekends with weekdays and high
precipitation with no precipitation. Additionally, line plots
visualizing distance, speed and duration per hour of the day comparing
weekends with weekdays were carried out. Cartographic maps visualizing
trajectory points of students comparing weekends with weekdays and
precipitation and no precipitation complement the analysis.

```{r}
#| label: load packages and clean data

# clear space 
rm(list=ls())

# load packages 
library("readr")
library("dplyr")
library("ggplot2")
library("sf")
library("terra")
library("tmap")
library("gitcreds")
library("dplyr")
library("SimilarityMeasures")
library("lubridate")
library("plotly")

# load clean data 
posmo <- read_delim("posmo_data/posmo_trips.csv")
```

# RESULTS

## Impact of the day of the week

### Spatial Analysis

A higher trip distance resulted for every travel mode on weekends
compared to weekdays. Generally, the distance travelled by foot (travel
mode: walk) had a lower mean than other travel modes. On the contrary it
can be seen, that the mean distance travelled by bike is in the same
range as the travel modes car and train.

```{r, fig.cap = "Figure 3: Trip distance per transport mode compared between weekends and weekdays, the trip distance is logarythmised"}
#| label: Boxplot comparing trip distance between weekend and weekday of travelmodes

  ggplot() +
  geom_boxplot(data = posmo, aes(day_week, log(trip_dis), fill = day_week)) +
  labs(title = "Trip Distance per Transport Mode", subtitle = "Comparing Weekend vs. Weekday", fill = "Day of the Week") +
  ylab("Log Trip Distance [m]") +
  xlab("Transport Mode") +
  scale_fill_manual(values = c("weekday" = "lavenderblush3", "weekend" = "aquamarine3"), labels = c( "Weekday", "Weekend")) +
    facet_wrap(~transport_mode, nrow = 1) +
    theme_minimal() +
  theme(axis.text.x=element_blank())
```

There is a tendency of trip speed being higher on the weekends than
during the week except for the travel mode train, where the relationship
is reversed. Again, we find the mean trip speed by bike being in a
similar range as when students travelled by car and train, whereas the
travel mode bus conforms to when students travelled by horse.

```{r, fig.cap = "Figure 4: Trip speed per transport mode compared between weekends and weekdays, the trip speed is logarythmised"}
#| label: Boxplot comparing trip speed between weekend and weekday of travelmodes

ggplot(posmo, aes(day_week, log(trip_speed), fill = day_week)) +
  geom_boxplot() +
  labs(title = "Trip Speed per Transport Mode", subtitle = "Comparing Weekend vs. Weekday", fill = "Day of the Week") +
  ylab("Log Speed [m/s]") +
  xlab("Transport Mode") +
  scale_fill_manual(values = c("weekday" = "lavenderblush3", "weekend" = "aquamarine3"), labels = c( "Weekday", "Weekend")) +
  facet_wrap(~transport_mode, nrow = 1) +
  theme_minimal() +
  theme(axis.text.x=element_blank())
```

### Temporal Analysis

The trip duration shows no clear trend, as the results appear
heterogeneous between weekends and weekdays for every travel mode
considered. There is however a tendency, that the mean trip duration is
lower when students are travelling by foot compared to all other travel
modes.

```{r, fig.cap = "Figure 5: Trip duration per transport mode compared between weekends and weekdays, the trip duration is logarythmised"}


#| label: Boxplot comparing trip duration between weekend and weekday of travelmodes

ggplot() +
  geom_boxplot(data = posmo, aes(day_week, log(trip_duration/60), fill = day_week)) +
   labs(title = "Trip Duration per Transport Mode", subtitle = "Comparing Weekend vs. Weekday", fill = "Day of the Week") +
  ylab("Log Duration [min]") +
  xlab("Transport Mode") +
  scale_fill_manual(values = c("weekday" = "lavenderblush3", "weekend" = "aquamarine3"), labels = c( "Weekday", "Weekend")) +
  facet_wrap(~transport_mode, nrow = 1) +
  theme_minimal() +
  theme(axis.text.x=element_blank())
```

### Cartographic Representation

```{r, fig.cap = "Figure 6: Cartographic representation of trajectory points from six students compared between weekends and weekdays"}
#| label: prepro and creation of visual map with convex hull

# transform posmo data into an sf object 
posmo <- st_as_sf(posmo, coords = c("X","Y"), crs = 2056) 

# 1. add grouping variable to the sf object
posmo_grouped <- group_by(posmo, dataset)

# 2. use summarise() to "dissolve" all point into a multipoint object
posmo_smry <- summarise(posmo_grouped)

# 3. run st_convex_hull()
mcp_posmo <- st_convex_hull(posmo_smry)


# set visual mode
tmap_mode("view")

# cartographic visualisation 
tm_shape(mcp_posmo) +
  tm_fill(col = "dataset", alpha = 0.4, title = "Student") +
  tm_shape(mcp_posmo) +
  tm_borders(col = "black") +
  tm_shape(posmo) +
  tm_dots(col = "day_week", title = "Day of the Week")
```

## Impact of the time of the day

```{r}
#| label: prepro metrics for plots

# load clean data 
posmo <- read_delim("posmo_data/posmo_trips.csv")

# round datetime to 1h 
posmo_round <- posmo |>
  mutate(hour = lubridate::hour(datetime))


# create dataframe, which calculates mean steplength, speed per hour over all dates comparing weekends and weekdays
posmo_day <- posmo_round |>
  group_by(hour, day_week)|>
  summarise(mean_dis = mean(trip_dis, na.rm = TRUE),
            mean_speed = mean(trip_speed, na.rm = TRUE),
            mean_duration = mean(trip_duration, na.rm = TRUE)) 
```

### Spatial Analysis

The trip distance per hour shows a higher trip distance on the weekends
compared to weekdays. After a peak of travelled distance at around 7 AM
on the weekend, a decreasing trend of distance per hour is shown, with
regular peaks (10 AM, 1 PM, 5 PM, 9 PM). The trip distance per hour
during the week shows a rather balanced pattern with regular peaks
during the day (5 AM, 7 AM, 9 AM, 1 PM, 5 PM, 8 PM) with the highest
peaks at around 1 PM, 5 PM and 8 PM.

```{r, fig.cap = "Figure 7: Mean trip distance for every hour of the day compared between weekdays and weekends"}

#| label: Mean trip distance per hour of the day compared between weekend/weekday

ggplot(posmo_day, aes(hour, mean_dis, col = day_week)) +
  geom_point() +
  geom_line(lwd = 0.7) +
  labs(title = "Mean Trip Distance over 24h", subtitle = "Compared between Weekend vs. Weekday", color = "Day of the Week") +
  scale_color_manual(values = c("weekday" = "mediumpurple4", "weekend" = "aquamarine3"), labels = c("Weekday", "Weekend")) +
  ylab("Mean Distance [m]") +
  xlab("Hour") +
  theme_minimal()
```

The trip speed shows a decreasing trend from around 4 AM to 11 PM on the
weekend. Additionally, trip speed per hour is mostly higher for weekends
compared to weekdays. On weekdays the trip speed per hour is relatively
stable during the day (7 AM to 4 PM) with two peaks at around 5 PM and 8
PM and a considerably lower speed at around 6 AM.

```{r, fig.cap = "Figure 8: Mean trip speed for every hour of the day compared between weekdays and weekends"}
#| label: Mean trip speed per hour of the day compared between weekend/weekday

ggplot(posmo_day, aes(hour, mean_speed, col = day_week)) +
  geom_point() +
  geom_line(lwd = 0.7) +
  labs(title = "Mean Speed over 24h", subtitle = "Compared between Weekend vs. Weekday", color = "Day of the Week") + 
  scale_color_manual(values = c("weekday" = "mediumpurple4", "weekend" = "aquamarine3"), labels = c("Weekday", "Weekend"))+
  ylab("Mean Speed [m/s]") +
  xlab("Hour") +
  theme_minimal()
```

### Temporal Analysis

The trip duration reaches its peak at around 4 AM on the weekends, while
on the weekdays the peak is at around 8 PM. Additionally, between 3 and
5 PM trip duration is considerably higher than the average on the
weekend, where during the week the trip duration is higher at around 8
AM and 5 PM compared to other hours of the day.

```{r, fig.cap = "Figure 9: Mean trip duration for every hour of the day compared between weekdays and weekends"}
#| label: Mean trip duration per hour of the day compared between weekend/weekday

ggplot(posmo_day, aes(hour, mean_duration/60, col = day_week)) +
  geom_point() +
    geom_line(lwd = 0.7) +
  labs(title = "Mean Trip Duration over 24h", subtitle = "Compared between Weekend vs. Weekday", color = "Day of the Week") +
  ylab("Mean Duration [min]") +
  xlab("Hour") +
  scale_color_manual(values = c("weekday" = "mediumpurple4", "weekend" = "aquamarine3"), labels = c("Weekday", "Weekend")) +
  theme_minimal()
```

## Impact of precipitation

### Spatial Analysis

Precipitation reduced trip distance by bike, foot, bus and car, while
the trip distance was higher for travel modes like train and tram. Rain
does not seem to have an effect on trip distance when travelling by
horse. Although, travel distances are impacted by precipitation, the
effect can be described as heterogeneous.

```{r, fig.cap = "Figure 10: Mean trip distance compared between precipiation and no/low precipitaion, where the trip distance is logarythmised"}

#| label: mean trip distance depending on precipitation

ggplot() +
  geom_boxplot(data = posmo, aes(rain_day, log(trip_dis), fill = rain_day)) +
    labs(title = "Trip Distance per Transport Mode", subtitle = "Comparing Precipitation vs. no Precipitation", fill = "Precipitation") +
  ylab("Log Distance [m]") +
  xlab("Transport Mode") +
  scale_fill_manual(values = c("rain" = "lavenderblush3", "no_rain" = "aquamarine3"), labels = c( "Precipitation", "No Precipitation")) +
  facet_wrap(~transport_mode, nrow = 1) +
  theme_minimal() +
  theme(axis.text.x=element_blank())
```

Trip speed at high precipitation is lower when travelling by bike, foot,
bus and car, while for travel modes such as train and tram speed was
higher. Travelling by horse does not seem to be effected by
precipitation. Although, travel speeds are impacted by precipitation,
the effect can be described as heterogeneous.

```{r, fig.cap = "Figure 11: Mean trip speed compared between precipiation and no/low precipitaion, where the trip speed is logarythmised"}
#| label: mean trip speed depending on precipitation

ggplot(posmo, aes(rain_day, log(trip_speed), fill = rain_day)) +
  geom_boxplot() +
   labs(title = "Trip Speed per Transport Mode", subtitle = "Comparing Precipitation vs. no Precipitation", fill = "Precipitation") +
  ylab("Log Speed [m/s]") +
  xlab("Transport Mode") +
  scale_fill_manual(values = c("rain" = "lavenderblush3", "no_rain" = "aquamarine3"), labels = c( "Precipitation", "No Precipitation")) +
  facet_wrap(~transport_mode, nrow = 1) +
  theme_minimal() +
  theme(axis.text.x=element_blank())
```

### Temporal Analysis

The duration of trips was shorter under heavy precipitation by bike, bus
and car, while students travelled longer by train and tram. Considering
travelling by horse and foot, precipitation does not seem to have an
effect on trip duration.

```{r, fig.cap= "Figure 12: Mean trip duration compared between precipiation and no/low precipitaion"}
#| label: mean trip duration depending on precipitation

ggplot() +
  geom_boxplot(data = posmo, aes(rain_day, trip_duration/60, fill = rain_day)) +
  labs(title = "Trip Duration per Transport Mode", subtitle = "Comparing Precipitation vs. no Precipitation", fill = "Precipitation") +
  ylab("Duration [min]") +
  xlab("Transport Mode") +
  scale_fill_manual(values = c("rain" = "lavenderblush3", "no_rain" = "aquamarine3"), labels = c( "Precipitation", "No Precipitation")) +
  facet_wrap(~transport_mode, nrow = 1) +
  theme_minimal() +
  theme(axis.text.x=element_blank())
```

### Cartographic Representation

```{r, fig.cap = "Figure 13: Cartographic representation of trajectory points from six students compared between precipitation and no/low precipitation. Additionally, all weather stations used are visualised as black points on the map."}
#| label: create visual map rain

# load weather stations 
legend <- read_delim("data/weather_legend.csv")

# create sf object of joined data
weather <- st_as_sf(legend, coords = c("E","N"), crs = 2056)

# transform posmo data into an sf object 
posmo <- st_as_sf(posmo, coords = c("X","Y"), crs = 2056) 

# 1. add grouping variable to the sf object
posmo_grouped <- group_by(posmo, dataset)

# 2. use summarise() to "dissolve" all point into a multipoint object
posmo_smry <- summarise(posmo_grouped)

# 3. run st_convex_hull()
mcp_posmo <- st_convex_hull(posmo_smry)

#choose map mode
tmap_mode("view")

# segmented visualisation 
tm_shape(mcp_posmo) +
  tm_fill(col = "dataset", alpha = 0.4, title = "Student") +
  tm_shape(mcp_posmo) +
  tm_borders(col = "black") +
  tm_shape(posmo) +
  tm_dots(col = "rain_day", title = "Precipitation") +
  tm_shape(weather) +
  tm_dots(title = "weather Stations")
```

# DISCUSSION

It is to be noted, that the investigated spatio-temporal movement
patterns of students cannot be generalized for the general population,
as it was shown that sociodemographics and travel behaviour of
university students differ from the general population, as well as
between students living on or off campus and students attending urban or
suburban campuses (Khattak et al., 2011). Furthermore, conclusions that
can be drawn by investigating our three environmental factor
individually are limited, as these factors interact with each other and
several additional factors lead to differences of time expenditure on
different travel modes such as for example temperature and gender.
Multi-channel sequence analysis (MCSA) are therefore particularly
relevant to study human mobility as they are able to simultaneously
consider multiple environmental variables (Brum-Bastos et al., 2018).

Additionally, during the course of this project it became clear, that
the Posmo App comes with severe uncertainties and tracking errors.
Quantifying those uncertainties would be subject to another semester
project in itself, which is why we assumed that the data we used was
correct and interpreted it as such.

## Patterns of the day of the week

The day of the week impacted spatio-temporal movement patterns of
students, where movement activity regarding trip distance and speed was
generally slightly higher on the weekends than during the week.

Looking at the use of bikes Sathishkumar et al. (2020) found, that most
bike trips during the week took around 24 minutes compared to 30 minutes
on the weekends, which aligns with our results, where bike trip duration
was slightly higher on weekends. Additionally, the use of public
transport during weekdays are only slightly higher in distance, speed
and sometimes lower in duration, which could indicate, that other
factors than only the weekday impact the spatio-temporal movement of
students over different travel modes. In line with this Brum-Bastos et
al. (2018) found, that rain has no key role on travel modes during the
weekend, but heavy rain decreases the use of public transport during the
week. Precipitation was a factor influencing 8 of 11 investigated days.

Since trip distance, duration and speed are interconnected parameters
and the pattern is clear for trip distance and gets weaker for speed and
becomes unrecognizable for trip duration, a reason could be our
imbalanced dataset. As we had to cut our datasets in time, to attain
seamless data over all days, from all individual participants, we ended
up with seven days of the data for weekdays and only four days for
weekends. This bias potentially has impacted our results. For further
studies we therefore suggest a larger number of participants and a
longer period of GPS-tracking, to achieve a more balanced dataset.

## Patterns of the time of the day

The time of the day impacted spatio-temporal movement patterns of
students compared between weekend and weekdays, where generally higher
movement activity per hour could be observed during weekends.
Spatio-temporal movement during the week was more balanced, with lower
peaks of distance, speed and duration during the day. In contrast, high
peaks of trip distance, speed and duration were found on the weekends,
early in the morning, with a general trend to decrease during the day,
often times with regular peaks.

Comparing our results to literature, the movement pattern Sathishkumar
et al. (2020) found, considering only the travel mode bike, stays in
contrast to our results found for the weekend, where there is the
shortest trip duration around 8 AM and the trips are the longest in the
afternoon and evening. However, in their study there was no distinction
between weekend and weekday, as well as only one travel mode was
investigated, making the comparison to our results difficult. Ryan et
al. (2010) and Brum-Bastos et al. (2018) found more varied activities on
weekends, which aligns with our findings of generally more peaks and
higher movement activity during the day on the weekends, which they
explain by having more scope for freedom of action on weekends compared
to external controls during the week. When looking at movement activity
during the day, it was shown that daylight impact movement behaviour of
humans, where less daylight results in less time walking, time spend on
public transport and more vehicular use during the weekdays. On weekends
this trend reverses, where walking and public transport become more
prominent during night hours, while vehicle use is more prominent during
daylight hours (Brum-Bastos et al., 2018). Additionally, more time at
home and less time shopping or socializing is spent during night hours
during the week while again, this relationship reverses on the weekends
(Brum-Bastos et al., 2018). This fits with our observations, as we found
a high peak of movement activity during 3 to 6 AM on the weekends, as
well as more and higher peaks of movement activity on the weekends.
During the weekdays the patterns of distance, speed and duration were
found to be more homogeneous with lower peaks, fitting our expectation
of an increase in activity of spatio-temporal movement behaviour during
commuting times (around 6 -- 8 AM and 3 -- 6 PM) . An additional peak
for distance, speed and duration at around 8 PM during the weekdays
could be explained by leisure activities in the evening.

## Precipiation Patterns

Precipitation impacted spatio-temporal movement patterns of students in
a hetergoeneous way, indicating no clear relationship between
precipitation and movement activity in general. However, some patterns
were emerging as the movement (distance, speed, and duration) of
students was reduced under heavy rain when travelling by bike, bus and
car, while it was increased when travelling by train and tram. These
results are in contrast to Brum-Bastos et al. (2018), where they found a
general trend of more vehicular use under heavy rain, while in our study
we found a general decrease in the use of cars.

It could be shown that speed and distance were lower under heavy rain
when travelling by foot, while the duration was not impacted by
precipitation patterns. In contrast to our finding, Brum-Bastos et al.
(2018) found an overall decrease in walking during rain. It was shown,
that when travel destinations are obligatory (e.g., work, lectures)
people change their travel mode more likely under heavy rain, meaning
that they drive to work instead of walk. However, if destinations are
linked to leisure activities, people more likely just postpone the task
instead of changing their travel mode (Connolly, 2008). Additionally,
during the week a decrease in public transport was observed under heavy
rain by Brum-Bastos et al. (2018). This finding are in contrast to our
results, where the use of some public transport means (train and tram)
was increased under heavy rain, although we did not include the
consideration of the days of the week.

However, it is to be said that those patterns found in our data are not
indicative, as we only had a sample size of 6 students and investigated
11 days, from which 8 days were considered having heavy rainfall.
Additionally, the results need to be interpreted with caution, as for
some trajectory points the weather stations are rather far away to serve
as an indication for precipitation (figure 13).

# SOURCES

## Data

-   **Trajectory data of students:** Posmo App
-   **Weather data:** https://gate.meteoswiss.ch/idaweb

## Literature

-   Brum-Bastos, V. S., Long, J. A., & Demšar, U. (2018). Weather
    effects on human mobility: A study using multi-channel sequence
    analysis. Computers, Environment and Urban Systems, 71, 131--152.
    https://doi.org/10.1016/j.compenvurbsys.2018.05.004

-   Bundesamt für Statistik. (2021). Pendlermobilität. Bundesamt für
    Statistik.
    https://www.bfs.admin.ch/bfs/de/home/statistiken/mobilitaet-verkehr/personenverkehr/pendlermobilitaet.html

-   Bundesamt für Statistik. (2023, March 28). Studierende an den
    universitären Hochschulen: Basistabellen - 1990-2022 \| Tabelle.
    Bundesamt für Statistik. https://www.bfs.admin.ch/asset/de/24345359

-   Connolly, M. (2008). Here Comes the Rain Again: Weather and the
    Intertemporal Substitution of Leisure. Source Journal of Labor
    Economics Journal of Labor Economics, 26(1), 73--100.
    https://doi.org/10.1086/522067

-   Guo, P., Sun, Y., Chen, Q., Li, J., & Liu, Z. (2022). The Impact of
    Rainfall on Urban Human Mobility from Taxi GPS Data. Sustainability,
    14(15), Article 15. https://doi.org/10.3390/su14159355

-   Kattiyapornpong, U., & Miller, K. E. (2009). Socio‐demographic
    constraints to travel behavior. International Journal of Culture,
    Tourism and Hospitality Research, 3(1), 81--94.
    https://doi.org/10.1108/17506180910940360

-   Khattak, A., Wang, X., Son, S., & Agnello, P. (2011). Travel by
    University Students in Virginia: Is This Travel Different from
    Travel by the General Population? Transportation Research Record:
    Journal of the Transportation Research Board, 2255(1).
    https://doi.org/10.3141/2255-15

-   Kyaing, K., Lwin, K., & Sekimoto, Y. (2017). Human mobility patterns
    for different regions in Myanmar based on CDRs data. IPTEK Journal
    of Proceedings Series, 3(6).

-   Liu, X., Sun, L., Sun, Q., & Gao, G. (2020). Spatial Variation of
    Taxi Demand Using GPS Trajectories and POI Data. Journal of Advanced
    Transportation, 2020, e7621576. https://doi.org/10.1155/2020/7621576

-   Medienmitteilung Kanton ZH. (2021). Bildung in Zahlen. Kanton
    Zürich.
    https://www.zh.ch/de/news-uebersicht/medienmitteilungen/2021/07/bildung-in-zahlen.html

-   MeteoSwiss. (2023). Precipitation. Federal Office of Meterology and
    Climatology MeteoSwiss. Retrieved 19 June 2023, from
    https://www.meteoschweiz.admin.ch/wetter/wetter-und-klima-von-a-bis-z/niederschlag.html

-   Ryan, R. M., Bernstein, J. H., & Brown, K. W. (2010). Weekends,
    Work, and Well-Being: Psychological Need Satisfactions and Day of
    the Week Effects on Mood, Vitality, and Physical Symptoms. Journal
    of Social and Clinical Psychology, 29(1), 95--122.
    https://doi.org/10.1521/jscp.2010.29.1.95

-   Sathishkumar, Cho, Y., & Jangwoo, Park. (2020). Seoul bike trip
    duration prediction using data mining techniques. IET Intelligent
    Transport Systems, 14(11), 1465--1474.
    https://doi.org/10.1049/iet-its.2019.0796
